Capon Beamforming and Moving Objects - An Analysis of Lateral Shift-Invariance

### 2014-02-20 HRUIofMovingObjects/response_to_reviewers.tex

As my scores indicate, I think the paper addresses an interesting problem and does it very well. Beside two VERY minor typos, I think it can be published unaltered.
1. "In" missing at the beginning of the Introduction section.
2. In section IIB, the opening sentence, there is an 's' missing in 'produce'.

\textit{We want to thank the reviewer for an encouraging review. Typos have been corrected.}

\section*{Reviewer: 2}

Thanks to the authors for addressing the problem of the temporal aspects in minimum variance beamforming, demonstrating the extra requirements on spatial sampling, and demonstrating how to efficiently oversample. The paper is good but I would like it to go through a second review round based on the following comments:
\textit{We want to thank the reviewer for an encouraging review. Below follows our response to the comments given.}
\\\\
Page2 column2 lines 34-35, and Fig. 1, L, M and d have not been introduced (need to have read the previous literature or equation (4) to know their meaning). I would suggest that in the introduction we do not need these details.
\textit{These details have now been removed from the introduction.}
\\\\
Equation (8) the two factors p and q are confusing. Because q is always one, I would suggest omitting it.  Let us say that the oversampling is used both for Capon and for reducing folding of frequencies over the bandwidth.
\textit{We do not find this confusing, $p$ is always 2 and $q$ is 4 and 16 when oversampling on transmit is applied for the Capon beamformer (as mention in the first paragraph of Section IV-A). In all the other examples $q=1$. The $p$ and $q$ is there to clearly separate oversampling from the Nyquist requirement.}
\\\\
At what depth are the LSV plots taken with respect to the focal depth? Does it matter?
\textit{The shift-invariance caused by under-sampling will be at its maximum at the transmit focus. Simulated LSV-plots are generated from this point, as specified on page 5, first paragraph. In the in-vitro data, the point scatterer closest to the transmit focus is selected.}
\\\\
Please specify the characteristics of the M5S probe (size, number of elements) if GE permits. Is it a phased or a linear array? What is the physical beam spacing, the field of view, the number of transmit and receive beams in the base (non oversampled) configuration? What is the transmit focal depth? Are transmit and receive apodization applied? Without this information, the work is not reproductible.
\textit{More extensive information about the probe and scan sequence have been added.}
\\\\
In section “IV-A. Oversampling in transmit”: please confirm (specify) that 1x1 receive beamforming is performed at this point.
\textit{We have now specified that our starting point is a system with a one-to-one relationship between tx and rx beams.}
\\\\
In section “IV-B. Oversampling by parallel receive beamforming”, could you explain in more details what are the differences between Capon + STB and the technique proposed by Rabinovitch (2013) – reference [27]?
\textit{There are no differences. Rabinovitch (2013), reference [25], does not propose any new technique, it is just Capon + STB.}
\\\\
Page10 Column1 line56  write “should be removed” instead of “should be remove”
\textit{Misspelling has been corrected.}
\\\\
Page 10 Column 1 lines 40-60, you state that with synthetic transmit beamforming, some of the artifacts seen in oversampling in receive by phase rotation could be eliminated. It would be interesting to see such technique implemented.
\textit{It would, but while minimizing the length of the manuscript an actual result was left out. It should, however, be clear from how STB corrects errors introduced by PRB, that something similar will work for phase-rotated rx-beams.}
\\\\
P10 column 2 line 28-29 write “too extreme” instead of “to extreme”
\textit{Misspelling has been corrected.}
\\\\
Fig 6a: specify which beamforming is applied here (I suspect delay and sum with no multiline receive?). Maybe you want to consider showing still images of Capon and Capon-oversampled as well (not necessarily on fig. 6), just in case some people won’t look at the videos.
\textit{We have now specified that the image is processed with delay-and-sum beamformer. Default values for $p$ and $q$ are used and are therefore not specified. Images are not added in order to minimize the length of the paper.}
\end{document}

### 2014-02-20 HRUIofMovingObjects/jrnl_article.tex
Ready for 2nd submit.

%%% My commands
\newcommand{\mat}[1]{\mathbf{#1}}
\renewcommand{\vec}[1]{\mathbf{#1}}
\newcommand{\R}{$\mat{\hat{R}}$ }

@@

\IEEEPARstart{I}{n} the last decade several authors have applied the Capon beamformer (also known as the minimum variance beamformer) for medical ultrasound imaging \cite{Synnevag2007, Vignon2008, Viola}. The method has been shown to improve lateral resolution and edge sharpness \cite{Synnevag2007, Synnevag2009, Chen2011} in ultrasound images. This is accomplished by adaptively adjusting the array weights based on element-to-element covariance in order to minimize interference. The method is, however, also highly sensitive to model errors \cite{Mestre2006, Widrow1982, Wax1996, Wax1996a}. This has led to the development of several robust versions of the Capon beamformer with e.g. spatial smoothing \cite{Shan1985} or diagonal loading \cite{JianLi2003} of the covariance estimate, or by applying steering vector uncertainty sets \cite{Lorenz2005, Rubsamen2013}. Spatial smoothing and diagonal loading, together with temporal smoothing of the covariance estimate, have been applied in earlier work on Capon beamforming for medical ultrasound imaging \cite{Synnevag2009}. The first two trade resolution for robustness and temporal smoothing helps to preserve speckle statistics for high resolution parameters \cite{Synnevag2007a}. Steering vector uncertainty sets, however, have not been applied for medical ultrasound imaging.

@@

In \cite{Cox1973}, Cox presents in great detail how the amount of mismatch between the selected steering vector and the signal propagation vector impacts the beamformer output in simple situations. He also states the following important fact about the Capon beamformer [the $\vec{k}_3$ processor]: 
\begin{quote}
"\textit{... a $\vec{k}_3$ processor requires more closely spaced beams than a ... $\vec{k}_1$-processor in order to avoid serious signal suppression effects being introduced on signals arriving from directions between the beams}", 
\end{quote}
where $\vec{k}_1$ is the delay-and-sum (DAS) beamformer. In Fig.\,\ref{fig:das_capon_beams} we have plotted the two-way response for DAS and the Capon beamformer using (26) and (32) in \cite{Cox1973}. A 64-element linear array with $\lambda/2$ element spacing, where $\lambda$ is the wavelength, is scanned over a sector equal to four times the array resolution using narrow band excitation. Located in the farfield of the sector is an object moving parallel to the array surface. The plot shows the high precision scanned output when the object is located at plus and minus half the required beam spacing for DAS [See (\ref{eq:resolution})]. %No spatially white noise is present. Hence, the interference-noise matrix ($\sigma_0^2\mat{Q}$ in \cite{Cox1973}) is the identity. 
The effect of typical subarray averaging and diagonal loading ($L=32$ and $d=1/100$) have been added to the Capon beamformer output. We see that a DAS beamformer, using the required beam spacing, could yield a $1.9$ dB scalloping loss \cite{Harris1978} if the object is moving parallel to the array. However, if the Capon beamformer is used with the same beam density the loss could be as large as $27$ dB. This imaging system is then highly sensitive to small spatial shifts, either by the object or the array itself. A $1.9$ dB scalloping loss for the Capon beamforming is obtained for this specific example when the beam spacing or object distance is decreased with a factor of 25.

\begin{figure}[!t]
\centerline{
\includegraphics[width=\linewidth]{\img fig1_2way.eps}
}
\caption{Estimated drop in output power mid-way between receive beams for delay-and-sum and Capon beamforming ($M=64$, $L=32$, $d=0.01$) applied in a farfield-narrowband setting. %The scanned output from both beamformers is plotted for two different time instances. For the first instances an object is located at an angle equal to half the required beam spacing for delay-and-sum, and for the latter the object is moved an angular distance corresponding to the system resolution. If the beam density was reduced to one beam per system resolution the maximum scalloping loss is found at zero degrees.
}
\label{fig:das_capon_beams}
\end{figure}

@@

Figure \ref{fig:das_shiftinvariant} presents such a LSV-plot for an imaging system using the DAS beamformer with a beam spacing equal to (\ref{eq:resolution}). Dotted vertical lines indicate position of transmit-receive beam pairs. The point scatterer response has been simulated using Field II \cite{Jensen1992, Jensen1996a} with a probe setup equal to the M5S probe by GE Vingmed Ultrasound, Horten, Norway. The probe was excited with a 2.5 MHz and 1.5 cycles long pulse. The transmit focus in azimuth and elevation were configured to be equal and the point scatterer was moved through this point. We see that the contours in Fig.\,\ref{fig:das_shiftinvariant} are approximately diagonal (low P-MAE) and that the amplitude variation among the maximum points is minimal (low A-MAE). We therefore conclude that the system is locally laterally shift-invariant. %In this, and the following LSV-plots based on simulations, the scan sequence always has a beam at zero degrees and the point scatterer is located in the center of this beam at one time instance.

@@

Figure \ref{fig:capon_oversampling_pr} shows the phase rotation method applied on the same scan sequence as in Fig\,\ref{fig:capon} with four and sixteen times oversampling respectively. Note that the figure is zoomed compared with Fig\,\ref{fig:capon}. As with oversampling on transmit there is some variation when the point moves in and out of focus. The maximum scalloping loss is therefore large. The extreme local variations are however reduced significantly. Comparing Fig\,\ref{fig:capon_oversampling} with Fig\,\ref{fig:capon_oversampling_pr}, the errors introduced by the phase rotation method are best seen in sidelobes.

@@

\section{In-vitro phantom data}\label{sec:res}
In Fig.\,\ref{fig:LSV_invitro} we present the result of applying the Capon beamformer on \textit{in-vitro} data with and without oversampling by phase rotation. The data was acquired using a modified Vivid E9 high-end cardiac ultrasound scanner by GE Vingmed Ultrasound, Horten, Norway, equipped with a M5S transducer operating in harmonic mode with a receive frequency of 3.7 MHz. The transmit beam density was set to match (\ref{eq:resolution}) based on this frequency. Note that in harmonic imaging the generated transmit beam at the harmonic frequency will be wider and have lower sidelobes than the beam resulting from transmitting at the harmonic frequency \cite{Fedewa2004}. The lateral sampling is therefore increased compared with our simulations. A tissue mimicking phantom containing wire targets and anechoic cysts was imaged, and the data was acquired while the probe was moved across the phantom manually. Fig.\,\ref{fig:image_das} shows the first frame out of 14 in the acquired loop. %Movies of the whole loop processed with DAS and Capon beamforming with and without ten times oversampling by phase rotation can be found in the supplementary material \multimedia{Videos}.

One point scatterer in Fig\,\ref{fig:image_das} has been selected [white box], and then used to generate LSV-plots. Figure \ref{fig:phantom_das} shows the resulting LSV-plot when using the DAS beamformer. The scalloping loss is less than 1 dB, and the system is by that locally laterally shift-invariant. As a consequence point scatterers have a consistent appearance when subjected to motion \multimedia{Media Movie 1}.

>>>>

%%% My commands
\newcommand{\mat}[1]{\mathbf{#1}}
\renewcommand{\vec}[1]{\mathbf{#1}}
\newcommand{\R}{$\mat{\hat{R}}$ }
\newcommand{\degree}{\ensuremath{^\circ}}

@@

\IEEEPARstart{I}{n} the last decade several authors have applied the Capon beamformer (also known as the minimum variance beamformer) for medical ultrasound imaging \cite{Synnevag2007, Vignon2008, Viola}. The method has been shown to improve lateral resolution and edge sharpness in ultrasound images \cite{Synnevag2007, Synnevag2009, Chen2011}. This is accomplished by adaptively adjusting the array weights based on element-to-element covariance in order to minimize interference. The method is, however, also highly sensitive to model errors \cite{Mestre2006, Widrow1982, Wax1996, Wax1996a}. This has led to the development of several robust versions of the Capon beamformer with e.g. spatial smoothing \cite{Shan1985} or diagonal loading \cite{JianLi2003} of the covariance estimate, or by applying steering vector uncertainty sets \cite{Lorenz2005, Rubsamen2013}. Spatial smoothing and diagonal loading, together with temporal smoothing of the covariance estimate, have been applied in earlier work on Capon beamforming for medical ultrasound imaging \cite{Synnevag2009}. The first two trade resolution for robustness and temporal smoothing helps to preserve speckle statistics for high resolution parameters \cite{Synnevag2007a}. Steering vector uncertainty sets, however, have not been applied for medical ultrasound imaging.

@@

In \cite{Cox1973}, Cox presents in great detail how the amount of mismatch between the selected steering vector and the signal propagation vector impacts the beamformer output in simple situations. He also states the following important fact about the Capon beamformer [the $\vec{k}_3$ processor]: 
\begin{quote}
"\textit{... a $\vec{k}_3$ processor requires more closely spaced beams than a ... $\vec{k}_1$-processor in order to avoid serious signal suppression effects being introduced on signals arriving from directions between the beams}", 
\end{quote}
where $\vec{k}_1$ is the delay-and-sum (DAS) beamformer. In Fig.\,\ref{fig:das_capon_beams} we have plotted the two-way response for DAS and the Capon beamformer using (26) and (32) in \cite{Cox1973}. A 64-element linear array with $\lambda/2$ element spacing, where $\lambda$ is the wavelength, is scanned over a sector equal to four times the array resolution using narrow band excitation. Located in the farfield of the sector is an object moving parallel to the array surface. The plot shows the high precision scanned output when the object is located at plus and minus half the required beam spacing for DAS [See (\ref{eq:resolution})]. %No spatially white noise is present. Hence, the interference-noise matrix ($\sigma_0^2\mat{Q}$ in \cite{Cox1973}) is the identity. 
The effect of typical subarray averaging and diagonal loading have been added to the Capon beamformer output. We see that a DAS beamformer, using the required beam spacing, could yield a $1.9$ dB scalloping loss \cite{Harris1978} if the object is moving parallel to the array. However, if the Capon beamformer is used with the same beam density the loss could be as large as $27$ dB. This imaging system is then highly sensitive to small spatial shifts, either by the object or the array itself. A $1.9$ dB scalloping loss for the Capon beamforming is obtained for this specific example when the beam spacing or object distance is decreased with a factor of 25.

\begin{figure}[!t]
\centerline{
\includegraphics[width=\linewidth]{\img fig1_2way.eps}
}
\caption{Estimated drop in output power mid-way between receive beams for delay-and-sum and Capon beamforming (with typical robustification parameters) applied in a farfield-narrowband setting. %The scanned output from both beamformers is plotted for two different time instances. For the first instances an object is located at an angle equal to half the required beam spacing for delay-and-sum, and for the latter the object is moved an angular distance corresponding to the system resolution. If the beam density was reduced to one beam per system resolution the maximum scalloping loss is found at zero degrees.
}
\label{fig:das_capon_beams}
\end{figure}

@@

Figure \ref{fig:das_shiftinvariant} presents such a LSV-plot for an imaging system using the DAS beamformer with a beam spacing equal to (\ref{eq:resolution}). Dotted vertical lines indicate position of transmit-receive beam pairs. The point scatterer response has been simulated using Field II \cite{Jensen1992, Jensen1996a} with a probe setup equal to the M5S-D phased array by GE Vingmed Ultrasound, Horten, Norway. The probe was excited with a 2.5 MHz and 1.5 cycles long pulse. The transmit focus in azimuth and elevation were configured to be equal and the point scatterer was moved through this point. We see that the contours in Fig.\,\ref{fig:das_shiftinvariant} are approximately diagonal (low P-MAE) and that the amplitude variation among the maximum points is minimal (low A-MAE). We therefore conclude that the system is locally laterally shift-invariant. %In this, and the following LSV-plots based on simulations, the scan sequence always has a beam at zero degrees and the point scatterer is located in the center of this beam at one time instance.

@@

Figure \ref{fig:capon_oversampling_pr} shows the phase rotation method applied on the same scan sequence as in Fig\,\ref{fig:capon} with four and sixteen times oversampling on receive respectively ($q$-factor). Note that the figure is zoomed compared with Fig\,\ref{fig:capon}. As with oversampling on transmit there is some variation when the point moves in and out of focus. The maximum scalloping loss is therefore large. The extreme local variations are however reduced significantly. Comparing Fig\,\ref{fig:capon_oversampling} with Fig\,\ref{fig:capon_oversampling_pr}, the errors introduced by the phase rotation method are best seen in sidelobes.

@@

\section{In-vitro phantom data}\label{sec:res}
In Fig.\,\ref{fig:LSV_invitro} we present the result of applying the Capon beamformer on \textit{in-vitro} data with and without oversampling by phase rotation. The data was acquired using a modified Vivid E9 high-end cardiac ultrasound scanner by GE Vingmed Ultrasound, Horten, Norway, equipped with a M5S-D phased transducer operating in harmonic mode at 3.34 MHz. The transmit beam density was set to match (\ref{eq:resolution}) based on this frequency, and on a 65\degree sector we therefore distributed 111 tx-beams. Transmit focus is located at 8 cm range. Note that in harmonic imaging the generated transmit beam at the harmonic frequency will be wider and have lower sidelobes than the beam resulting from transmitting at the harmonic frequency \cite{Fedewa2004}. The lateral sampling is therefore increased compared with our simulations. A tissue mimicking phantom containing wire targets and anechoic cysts was imaged, and the data was acquired while the probe was moved across the phantom manually. Fig.\,\ref{fig:image_das} shows the first frame out of 14 in the acquired loop processed with DAS beamforming. %Movies of the whole loop processed with DAS and Capon beamforming with and without ten times oversampling by phase rotation can be found in the supplementary material \multimedia{Videos}.

One point scatterer in Fig\,\ref{fig:image_das} has been selected [white box], and then used to generate LSV-plots. The point scatterer is located close to the transmit focus. Figure \ref{fig:phantom_das} shows the resulting LSV-plot when using the DAS beamformer. The scalloping loss is less than 1 dB, and the system is by that locally laterally shift-invariant. As a consequence point scatterers have a consistent appearance when subjected to motion \multimedia{Media Movie 1}.
